---
title: "Oligopoly 2020 - Chapter 1 - Data preparation"
author: "Philippe Mongeon"
date: "01/05/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: true
    theme: cosmo  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
    css: my.css   # you can add your custom css, should be in same folder
---

```{css, echo=FALSE}
h1 {
  padding-top: 30px;
}
h2 {
  padding-top: 20px;
}
h3{
  padding-top:10px;
}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


list_of_packages <- c("tidyverse","sqldf","readxl","flextable","kableExtra")
lapply(list_of_packages, library, character.only = TRUE)
remove(list_of_packages)



```

# Data sources
## journals_wos.xlsx
* **year**: year of publication of the paper
* **source_id**: CWTS unique ID for the journal
* **source_title**: CWTS journal name
* **issn_e**: electronic ISSN of the journal
* **issn_print**: print ISSN of the journal
* **issn**: ISSN obtained from the raw WoS data
* **source_title**: journal name in the raw WoS data
* **publisher**: publisher obtained from the raw WoS data
* **n_pubs**: number of distinct publications with DOI for the combination of all other columns

Even though it makes the file a bit messy, We keep the different journal names and ISSNs to improve the matching with the APC data.

## apcs.csv

This is a file that combines all the data from Lisa's GitHub (https://github.com/lmatthia/publisher-oa-portfolios).  
These columns are from the original files on the GitHub

* **issn**: ISSN of the journal:
* **journal_title**:
* **oa_model**:
* **apc**:
* **year**: year of the data collection  
  
These columns have been added to the original data

* **currency**: currency of the APC
* **apc_member**: APC for members extracted from the original APC column.
* **apc_non_member**: APC for non-members extracted from the original APC column
* **filename**: name of the GitHub file where the entry came from.

```{r read_data, echo = FALSE}
cambridge <- read.csv("C:/Users/phil_/Documents/GitHub/cost-of-oa/publisher-oa-portfolios-master/cambridge_apcs.csv", sep=";") %>% 
  mutate(filename = "cambridge_apcs.csv", currency="USD", publisher="Cambridge") %>% 
  rename(issn = ï..issn)
copernicus <- read.csv("C:/Users/phil_/Documents/GitHub/cost-of-oa/publisher-oa-portfolios-master/copernicus_apcs.csv", sep=";") %>% 
  mutate(filename = "copernicus_apcs.csv", currency="USD", publisher="Copernicus") %>% 
  rename(issn = ï..issn, apc = apc_in_EUR)
elsevier <- read.csv("C:/Users/phil_/Documents/GitHub/cost-of-oa/publisher-oa-portfolios-master/elsevier_oa_and_hybrid.csv", sep=";") %>% 
  mutate(filename ="elsevier_oa_and_hybrid.csv", currency="EUR", publisher = "Elsevier") %>% 
  rename(issn = ï..issn)
hindawi <- read.csv("C:/Users/phil_/Documents/GitHub/cost-of-oa/publisher-oa-portfolios-master/hindawi_apcs.csv", sep=";") %>% 
  mutate(filename ="hindawi_apcs.csv", currency="USD", publisher = "Hindawi") %>% 
  rename(issn = ï..issn, oa_model = oa_mode) %>% 
  mutate(apc = as.factor(apc))
nature <- read.csv("C:/Users/phil_/Documents/GitHub/cost-of-oa/publisher-oa-portfolios-master/nature_oa_and_hybrid.csv", sep=";") %>% 
  mutate(filename ="nature_oa_and_hybrid.csv", currency="USD", publisher = "Nature") %>% 
  rename(issn = ï..issn, oa_model = oa_mode) %>% 
  mutate(apc = as.factor(apc))
oup <- read.csv("C:/Users/phil_/Documents/GitHub/cost-of-oa/publisher-oa-portfolios-master/oup_apcs.csv", sep=";") %>% 
  mutate(filename ="oup_apcs.csv", currency="USD", publisher = "OUP") %>% 
  rename(issn = ï..issn, oa_model = oa_mode)
sage <- read.csv("C:/Users/phil_/Documents/GitHub/cost-of-oa/publisher-oa-portfolios-master/sage_oa_and_hybrid.csv", sep=";") %>% 
  mutate(filename ="sage_oa_and_hybrid.csv", currency="USD", publisher="Sage") %>% 
  rename(issn = ï..issn, oa_model = oa_mode) %>% 
  select(issn, journal_title, oa_model, apc, year, filename, currency, publisher)
springer_nature <- read.csv("C:/Users/phil_/Documents/GitHub/cost-of-oa/publisher-oa-portfolios-master/springer_nature_oa_and_hybrid.csv", sep=";") %>% 
  mutate(filename ="springer_nature_oa_and_hybrid.csv", currency="USD", publisher = "Springer-Nature") %>% 
  rename(issn = ï..issn)
springer <- read.csv("C:/Users/phil_/Documents/GitHub/cost-of-oa/publisher-oa-portfolios-master/springer_oa_and_hybrid.csv", sep=";") %>% 
  mutate(filename ="springer_oa_and_hybrid.csv", currency="USD", publisher = "Springer") %>% 
  rename(oa_model = oa_mode) %>% 
  mutate(apc = as.factor(apc))
taylor_francis <- read.csv("C:/Users/phil_/Documents/GitHub/cost-of-oa/publisher-oa-portfolios-master/taylor_francis_oa_and_hybrid.csv", sep=";") %>% 
  mutate(filename ="taylor_francis_oa_and_hybrid.csv", currency="USD", publisher = "Taylor & Francis") %>% 
  rename(issn = ï..issn, oa_model = oa_mode)
wiley <- read.csv("C:/Users/phil_/Documents/GitHub/cost-of-oa/publisher-oa-portfolios-master/wiley_oa_and_hybrid.csv", sep=";") %>% 
  mutate(filename ="wiley_oa_and_hybrid.csv", currency="USD", publisher = "Wiley") %>% 
  rename(issn = ï..issn) %>% 
  select(issn, journal_title, oa_model, apc, year, filename, currency, publisher)


# Merge into a single file
apc <- do.call("rbind",list(cambridge,copernicus,elsevier,hindawi,nature,oup,sage,springer,springer_nature,taylor_francis,wiley)) %>% 
  mutate(journal_tile = tolower(journal_title)) %>% 
  mutate(issn = tolower(issn),journal_title = tolower(journal_title))

#remove the original files
obj <- ls()
remove(list=obj[-1])
remove(obj)

# Pivot

total <- apc %>% 
  subset(year > 2009) %>% 
  select(publisher, journal_title) %>% 
  group_by(publisher) %>% 
  summarize(unique_journals = n_distinct(journal_title)) %>% 
  ungroup
```

  \
  <div style="text-align:center;">**Table 1. Summary of Lisa's data**</div>

```{r, echo = FALSE}
apc %>%
  subset(year > 2009) %>%
  select(publisher, year) %>%
  group_by(publisher,year) %>%
  summarize(nb_journal = n()) %>%
  ungroup %>%
  drop_na(year) %>%
  arrange(year) %>%
  mutate(nb_journal = if_else(is.na(nb_journal), "", as.character(nb_journal))) %>% 
  pivot_wider(names_from=year, values_from=nb_journal) %>%
  left_join(total, by="publisher") %>%
  flextable(col_keys = c("publisher", seq(from = 2010, to = 2020, by = 1), "unique_journals")) %>%
  bold(part="header", bold = TRUE) %>%
  bold(j=1) %>%
  autofit() %>%
  align(part="header", align = "center") %>%
  align(part="body", align = "center") %>%
  fix_border_issues()
```

  \
  \
And this is a HTML table made with the *Kable* package  
  \

```{r html table, echo=FALSE, warning=FALSE}

options(knitr.kable.NA = 0)

apc %>% 
  subset(year > 2009) %>% 
  select(publisher, year) %>% 
  group_by(publisher,year) %>% 
  summarize(nb_journal = n()) %>% 
  ungroup %>% 
  drop_na(year) %>% 
  arrange(year) %>% 
  pivot_wider(names_from=year, values_from=nb_journal) %>% 
  left_join(total, by="publisher") %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position = "center") 

remove(total)

```

# Publisher data (2010-)

## Raw publisher list
This first step constited in extracting the raw publisher information from web of science articles and reviews published since 2010. This data contais **2,603** publishers that collectively published **14,182,223** articles or review. The code used to collect it from the CWTS database using the following code, and the file can be found in https://github.com/pmongeon/cost-of-oa/data/wos_publisher_raw.csv

```{sql eval=FALSE, echo=TRUE}

Select a.pub_year, a.source_id, e.source_title, d.pu as publisher, count(distinct a.ut) as n_pubs
from wos_2013..pub a
left join wosdb1952..ut b on b.ut = a.ut
left join wosdb1952..ui c on c.ui = b.ui
left join wosdb1952..PU d on d.UI = c.ui
left join wos_2013..source e on e.source_id = a.source_id
where a.pub_year >= 2010 and a.doc_type_id < 3 
group by a.pub_year, e.source_title, d.PU, a.source_id
```

## Disambiguating publishers

Leigh-Ann then created a conversion table to unify the different charachter strings referring to a same publisher under a unique name. 
It should be noted that journals published by specific university departments were kept seperate and now merged under a single university as publisher.
The resulting conversion table contains the original name from teh raw data ("publisher_original" column) and the cleaned name ("publisher_clean" colum). The file can be access here: https://github.com/pmongeon/cost-of-oa/data/wos_publisher_raw.csv

## Identifying journals transition to and from large publishers

``` {r echo=FALSE}



data <- read_excel("C:/Users/phil_/Documents/GitHub/cost-of-oa/data/wos_journals.xlsx") %>% 
  arrange(pub_year)

title_list <- sqldf("
          select distinct source_id, lower(source_title) as title from data where source_title is not null
          ")
title <- title_list %>% 
  group_by(source_id) %>% 
  mutate(title = paste(title, collapse = "; ")) %>% 
  ungroup() %>% 
  select(source_id, title) %>% 
  unique()  

issn_list <- sqldf("
          select distinct source_id, issn from data where issn is not null
           ")
issn <- issn_list %>% 
  group_by(source_id) %>% 
  mutate(issn = paste(issn, collapse = "; ")) %>% 
  ungroup() %>% 
  select(source_id, issn) %>% 
  unique()  
    

publisher_conversion <- read.csv("C:/Users/phil_/Documents/GitHub/cost-of-oa/data/cleaned.csv") %>% 
  mutate(publisher=tolower(ï..publisher_do_not_modify),publisher_clean = tolower(publisher_clean)) %>% 
  select(publisher, publisher_clean)

publisher <- data %>% 
  select(source_id, pub_year, publisher) %>% 
  mutate(publisher = tolower(publisher)) %>% 
  left_join(publisher_conversion, by="publisher") %>% 
  mutate(publisher = publisher_clean) %>% 
  select(source_id, pub_year, publisher) %>% 
  unique() %>% 
  group_by(source_id, publisher) %>% 
  drop_na() %>% 
  mutate(min_year = as.character(min(pub_year))) %>% 
  mutate(max_year = as.character(max(pub_year))) %>%
  mutate(wos_publisher = paste(publisher," (",min_year,"-",max_year,")")) %>%
  ungroup() %>% 
  select(source_id, wos_publisher) %>% 
  unique() %>% 
  group_by(source_id) %>% 
  mutate(publisher = paste(wos_publisher, collapse = "; ")) %>% 
  ungroup() %>% 
  select(source_id, publisher) %>% 
  drop_na() %>% 
  unique() %>% 
  left_join(title, by="source_id") %>% 
  left_join(issn, by="source_id") %>% 
  unique() %>% 
  select(source_id, issn, title, publisher)

# match with Lisa's lists

match_issn <-issn_list %>%
   inner_join(select(apc, issn, year, publisher), by = "issn") %>%
  select(source_id, publisher, year) %>% 
   unique()

match_title <-title_list %>%
   inner_join(select(apc, journal_title, year, publisher), by = c("title" = "journal_title")) %>%
  select(source_id, publisher, year) %>% 
   unique()

publisher_lisa <- rbind(match_issn, match_title) %>% 
  arrange(year) %>% 
  unique() %>% 
  group_by(source_id, publisher) %>% 
  mutate(min_year = min(year), max_year=max(year)) %>% 
  mutate(publisher_lisa = paste(publisher," (",min_year,"-",max_year,")")) %>% 
  ungroup %>% 
  select(source_id, publisher_lisa) %>% 
  unique() %>% 
  group_by(source_id) %>% 
  mutate(publisher_lisa = paste(publisher_lisa, collapse = "; "))

# create final table

journal_list <- publisher %>%
  left_join(publisher_lisa, by="source_id") %>% 
  mutate(publisher_lisa = tolower(publisher_lisa))


big_5 <- c("springer","elsevier","taylor","wiley","sage")


a <- journal_list[unlist(lapply(big_5, function(x) grep(x, journal_list$publisher, fixed = TRUE))),]
b <- journal_list[unlist(lapply(big_5, function(x) grep(x, journal_list$publisher_lisa, fixed = TRUE))),]
c <- rbind(a,b) %>% unique() %>% 
  mutate(big_5 = "1") %>% 
  select(source_id, big_5)

publisher_list_tmp <- data %>% 
  select(source_id, pub_year, publisher) %>% 
  mutate(publisher = tolower(publisher)) %>% 
  left_join(publisher_conversion, by="publisher") %>% 
  mutate(publisher = publisher_clean) %>% 
  select(source_id, pub_year, publisher) %>% 
  unique() %>% 
  select(source_id, publisher) %>% 
  drop_na()

d <- publisher_list_tmp[unlist(lapply(big_5, function(x) grep(x, publisher_list_tmp$publisher, fixed = TRUE))),] %>% 
  mutate(big_5 = 1) %>% 
  select(source_id, publisher, big_5) %>% 
  unique()

publisher_list <- publisher_list_tmp %>% 
  left_join(d, by = c("source_id" = "source_id", "publisher" = "publisher")) %>% 
  mutate(other_publisher = ifelse(is.na(big_5),1,0)) %>% 
  select(source_id, other_publisher) %>% 
  unique() %>% 
  filter(other_publisher==1)

journal_list <- journal_list %>% 
  left_join(c, by="source_id") %>% 
  left_join(publisher_list, by="source_id")
  

write.csv(journal_list, "C:/Users/phil_/Documents/GitHub/cost-of-oa/data/journal_List_ost.csv", row.names = FALSE)

```


# Estimating the APCs paid to publishers on a given year

## Getting APC data for the journal

(Coming soon)

## Geting the OA status of publications with Unpaywall

(Coming soon)





