---
title: "os"
---

```{r}
library(RPostgres)
library(tidyverse)
{
  if(exists("db")) { 
    dbDisconnect(db) 
    rm(db)
  }
  db = dbConnect(Postgres(),
                 user = read.csv("db_connection.csv")$user,
                 password = read.csv("db_connection.csv")$pw,
                 host = read.csv("db_connection.csv")$host,
                 port = read.csv("db_connection.csv")$port,
                 db = "postgres",
                 options="-c search_path=projectdb_open_science_map")
}

x<-dbGetQuery(db, str_c("SELECT distinct wk.work_id, k.keyword, k.taxonomy_id, is_core
                     FROM keywords k
                     JOIN works_keywords wk ON wk.keyword_id = k.id
                     JOIN works w on w.id = wk.work_id
                     WHERE k.os_related = TRUE and w.is_core = FALSE"))


```{r}
data <- dbGetQuery(db, str_c("SELECT n.cluster, w.id, w.journal, w.pub_year, w.is_core, k.keyword, k.os_related, t.label as taxonomy
                          FROM works w
                          LEFT JOIN works_keywords wk ON wk.work_id = w.id
                          LEFT JOIN keywords k ON k.id = wk.keyword_id
                          LEFT JOIN taxonomy t ON t.id = k.taxonomy_id
                          INNER JOIN clusters_expanded n ON n.id = w.id
                          WHERE n.network = 'cc_dc'"))

write_csv(data, "data.csv")
```

## Number of papers

```{r}
n_papers <- data %>% 
  select(cluster, id) %>% 
  unique() %>% 
  group_by(cluster) %>% 
  reframe(n_papers = n())


data <- data %>% 
  filter(cluster %in% filter(n_papers, n_papers > 99)$cluster)

n_papers %>%
  arrange(desc(n_papers))
  kbl() %>% 
  kable_classic()
```

## Number of core papers

```{r}
library(kableExtra)
n_core_papers <- data %>%
  filter(is_core == T) %>% 
  select(cluster, id) %>% 
  unique() %>% 
  group_by(cluster) %>% 
  reframe(n_core_papers = n())

n_core_papers %>%
  arrange(desc(n_core_papers))
  kbl() %>% 
  kable_classic()


```

## Journals

```{r}
journal_freq <- data %>% 
  select(cluster, journal, id) %>% 
  unique() %>% 
  group_by(cluster, journal) %>% 
  reframe(freq = n()) %>% 
  group_by(cluster) %>% 
  mutate(rel_freq = freq/sum(freq)) %>% 
  group_by(journal) %>% 
  mutate(rel_freq_journal = freq/sum(freq)) %>% 
  ungroup() %>% 
  DT::datatable(filter = "top", rownames = F)

journal_freq

journal_freq$x$data %>% 
  write_csv("journal.csv")
```

## keywords

```{r}
keywords <- data %>% 
  select(cluster, id, keyword) %>% 
  unique() %>% 
  group_by(cluster, keyword) %>% 
  reframe(freq = n()) %>% 
  group_by(cluster) %>% 
  mutate(rel_freq = round(freq/sum(freq),3)) %>% 
  group_by(keyword) %>% 
  mutate(rel_freq_keyword = round(freq/sum(freq),3)) %>% 
  ungroup() %>% 
  DT::datatable(filter = "top", rownames = F)

keywords

keywords$x$data %>% 
  write_csv("keywords.csv")
```

## Open science related keywords

```{r}
keywords_os <- data %>% 
  filter(os_related == T) %>% 
  select(cluster, id, keyword) %>% 
  unique() %>%   
  group_by(cluster, keyword) %>% 
  reframe(freq = n()) %>% 
  group_by(cluster) %>% 
  mutate(rel_freq = round(freq/sum(freq),3)) %>% 
  group_by(keyword) %>% 
  mutate(rel_freq_keyword = round(freq/sum(freq),3)) %>% 
  ungroup() %>% 
  DT::datatable(filter = "top", rownames = F)

keywords_os

keywords_os$x$data %>% 
  write_csv("keyword_os.csv")
```

## Taxonomy

```{r}
taxonomy <- data %>% 
  filter(os_related == T) %>% 
  select(cluster, id, keyword, taxonomy) %>% 
  unique() %>% 
  group_by(cluster, taxonomy) %>% 
  reframe(freq = n()) %>% 
  group_by(cluster) %>% 
  mutate(rel_freq = round(freq/sum(freq),3)) %>% 
  group_by(taxonomy) %>% 
  mutate(rel_freq_taxonomy = round(freq/sum(freq),3)) %>% 
  ungroup() %>% 
  bind_rows(data %>% 
              filter(os_related == T) %>% 
              select(id, keyword, taxonomy) %>% 
              unique() %>% 
              group_by(taxonomy) %>% 
              reframe(freq = n()) %>% 
              mutate(rel_freq = round(freq/sum(freq),3)) %>% 
              group_by(taxonomy) %>% 
              mutate(rel_freq_taxonomy = round(freq/sum(freq),3)) %>% 
              mutate(cluster = 0) %>% 
              ungroup()) %>% 
  DT::datatable(filter="top", rownames = F)

taxonomy

taxonomy$x$data %>% 
  write_csv("taxonomy.csv")
```

# Temporal trends

## Clusters

```{r}
trend_cluster_n_papers<-data %>% 
  select(cluster, pub_year, id) %>% 
  unique() %>% 
  group_by(cluster, pub_year) %>% 
  reframe(freq = n()) %>% 
  mutate(rel_freq = freq/sum(freq)) %>% 
  bind_rows(data %>% 
              select(pub_year, id) %>% 
              unique() %>% 
              group_by(pub_year) %>% 
              reframe(freq = n()) %>% 
              mutate(rel_freq = freq/sum(freq)) %>% 
              mutate(cluster = 0)) %>% 
  ggplot() +
  aes(x=pub_year, y=freq) + 
  geom_line() +
  facet_wrap(facets = "cluster") +
  theme_classic() +
  xlab("Publication year") +
  ylab("Number of publications")

ggsave("trend_clusters_n_papers.tiff", device = "tiff", path = "tables and figures/", dpi = 300, units = "in", width = 8, height = 6)

trend_cluster_n_papers$data %>% 
  write_csv("trend_clusters_n_papers_data.csv")
```

## Taxonomy

```{r}
trend_taxonomy_n_papers <- data %>% 
  select(taxonomy, pub_year, id) %>%
  unique() %>% 
  drop_na() %>% 
  group_by(taxonomy, pub_year) %>% 
  reframe(freq = n()) %>% 
  mutate(rel_freq = freq/sum(freq)) %>% 
  ggplot() +
  aes(x=pub_year, y=freq) + 
  geom_line() +
  facet_wrap(facets = "taxonomy") +
  theme_classic() +
  xlab("Publication year") +
  ylab("Number of publications")

trend_taxonomy_n_papers

ggsave("trend_taxonomy_n_papers.tiff", device = "tiff", path = "tables and figures/", dpi = 300, units = "in", width = 8, height = 6)

trend_cluster_n_papers$data %>% 
  write_csv("trend_taxonomy_n_papers_data.csv")
```

